BeeModeling

### 2022-10-13 Fowler-GLOBI Modeling/GLOBI-Fowler Data Prep-reformatted.Rmd
removing interactions not at species or genus level

library(lifecycle)
library(rvest)
library(dplyr)
library(plotly)
library(tidyverse)
library(data.table)
library(readxl)
library(Taxonstand)

@@

```{r, echo = False}
options(max.print=1000000)
interactions_b <- read.csv('modeling_data/interactions_dup.csv')
ms <- read_xlsx('modeling_data/DATA_pollen-diet-breadth_list_12-5-21_ALR.xlsx', sheet = 3) %>%
  rename_all(tolower) %>%
  mutate(scientificName = paste0(genus, ' ',sub('.*\\. ','',species))) %>%
  select(scientificName,everything()) %>%
  rename(diet_breadth=`redefined pollen diet breadth`)
fowler <- read_csv("modeling_data/fowler_hostplants.csv")

#who are the specialsits?
#let's say specialist the MS list and everything on fowler list that's missing from ms list
ms_oligo = ms$scientificName[ms$diet_breadth %in% c('Oligolectic','Monolectic')]
add_from_fowler = unique(fowler$scientificName[!fowler$scientificName %in% ms$scientificName])
specialists = c(ms_oligo,add_from_fowler)

@@

# for (c in colnames(interactions_b)) {
#   interactions_b[[c]] <-  tolower(interactions_b[[c]])
# }
```

```{r}
#how many species in the missouri state data are in the globi data?
head(interactions_b)
n_distinct(interactions_b$sourceTaxonName[interactions_b$sourceTaxonName  %in% tolower(ms$scientificName)]) #800 species in the ms data

#how many are specialists
ms_specialists = ms %>% filter(diet_breadth != 'Polylectic')
n_distinct(interactions_b$sourceTaxonName[interactions_b$sourceTaxonName  %in% tolower(ms_specialists$scientificName)]) #503 specialist species

#how many are generalists
ms_generalists = ms %>% filter(diet_breadth == 'Polylectic')
n_distinct(interactions_b$sourceTaxonName[interactions_b$sourceTaxonName  %in% tolower(ms_generalists$scientificName)]) #296 generalist species

@@

interactions_reformat = interactions_b %>% 
  mutate(scientificName = capitalize(sourceTaxonSpeciesName), 
         bee_family = capitalize(sourceTaxonFamilyName),
         plant_family = capitalize(targetTaxonFamilyName), 
         plant_genus = capitalize(targetTaxonGenusName),
         plant_species = capitalize(targetTaxonName)) %>%
  filter(scientificName != '' & plant_family !='' & plant_genus != "")

#

@@

#add accepted_name column to plant_df
interactions_updated = interactions_reformat %>% 
  group_by(scientificName,bee_family,plant_species,plant_family) %>%
  summarize(n=n()) %>%
  ungroup %>%
  left_join(tpl_formatted) %>%
  select(-plant_species) %>%
  rename(plant_species = accepted_name) %>%
  mutate(plant_genus = sub(" .*","",plant_species))

@@

globi_summ = interactions_updated %>%
  group_by(scientificName, plant_genus, plant_family,bee_family) %>%
  summarize(n_genus = sum(n)) %>% ungroup()

@@

```{r}
#calculate the number of plant families each bee species is observed visiting
total_obs = globi_summ %>% group_by(scientificName) %>% summarize(n=sum(n_genus))

# degree_by_fam 
degree_by_fam = globi_summ %>% group_by(scientificName,plant_family) %>%
  summarize(n_family = sum(n_genus)) %>%
  left_join(total_obs) %>%
  mutate(freq_squared = (n_family/n)^2) %>% group_by(scientificName,n) %>%
  summarize(degree_family = n_distinct(plant_family), 
            degree5_family = sum(n_family >= 5), weighted_degree_fam = sum(freq_squared))# %>%

@@

globi_degree = degree_by_fam %>% left_join(degree_by_genus) %>%
  mutate(diet_breadth = ifelse(scientificName %in% specialists,'specialist','generalist'))
getwd()
# write_csv(globi_degree,'modelign_data/globi_degree.csv')

>>>>

library(lifecycle)
library(rvest)
library(dplyr)
library(plotly)
library(tidyverse)
library(data.table)
library(readxl)
library(Taxonstand)
library(vroom)

@@

```{r, echo = False}
options(max.print=1000000)
interactions_b <- vroom('modeling_data/interactions_dup.csv')
fowler <- read_csv("modeling_data/fowler_hostplants.csv")

@@

@@

interactions_reformat = interactions_b %>% 
  mutate(scientificName = capitalize(sourceTaxonSpeciesName), 
         bee_family = capitalize(sourceTaxonFamilyName),
         plant_family = capitalize(targetTaxonFamilyName), 
         plant_genus = capitalize(targetTaxonGenusName),
         plant_species = capitalize(targetTaxonName)) %>%
  filter(scientificName != '' & plant_family !='' & plant_genus != "")%>%
   filter(!targetTaxonRank %in% c('species',"genus"))

#
str(interactions_reformat)

@@

#add accepted_name column to plant_df
interactions_updated = interactions_reformat %>% 
  group_by(scientificName,bee_family,plant_species,plant_family) %>%
  summarize(n=n(),n_source_citations = n_distinct(sourceCitation),n_ref_citations = n_distinct(referenceCitation)) %>%
  ungroup %>%
  left_join(tpl_formatted) %>%
  select(-plant_species) %>%
  rename(plant_species = accepted_name) %>%
  mutate(plant_genus = sub(" .*","",plant_species))%>%
  mutate(plant_genus = ifelse(plant_genus == "Ipomaea","Ipomoea",plant_genus))

head(interactions_updated)

#let's also reformat so that the data re not grouped (each row a unique record)
nrow(interactions_reformat %>% left_join(tpl_formatted)) == nrow(interactions_reformat)

str(interactions_reformat)
globi_names = colnames(interactions_reformat)
globi_names[grepl('Sex',globi_names)]

each_interaction = interactions_reformat %>% 
  select(scientificName,bee_family,plant_species,plant_family,interactionTypeName,sourceTaxonRank, targetTaxonRank,referenceCitation,sourceCitation,decimalLongitude,decimalLatitude,sourceSexId,sourceSexName,targetSexId,targetSexName) %>%
  left_join(tpl_formatted) %>%
  select(-plant_species) %>%
  rename(plant_species = accepted_name) %>%
  mutate(plant_genus = sub(" .*","",plant_species))%>%
  mutate(plant_genus = ifelse(plant_genus == "Ipomaea","Ipomoea",plant_genus))

interactions_updated %>%
  select(scientificName,n,n_source_citations,n_ref_citations,plant_species) %>%
  filter(n>5)

# write_csv(interactions_reformat %>% filter(scientificName == "Agapostemon angelicus" & plant_species =="Helianthus annuus") %>% distinct(sourceCitation,referenceCitation),'look_at_refs.csv')

# write_csv(interactions_updated,'modeling_data/globi_names_updated.csv')
# write_csv(each_interaction,'modeling_data/globi_occ_names_updated.csv')
```

@@

globi_summ = interactions_updated %>%
  group_by(scientificName, plant_genus, plant_family,bee_family) %>%
  summarize(n_genus = sum(n),genus_source_citations=sum(n_source_citations)) %>% #this treats interactions of different plant species within the same genus as separate citations
  ungroup()

@@

```{r}

#calculate the number of plant families each bee species is observed visiting
total_obs = globi_summ %>% group_by(scientificName) %>% summarize(n=sum(n_genus))

# degree_by_fam 
degree_by_fam = globi_summ %>% group_by(scientificName,plant_family) %>%
  summarize(n_family = sum(n_genus),n_source_citations=sum(genus_source_citations)) %>% # %>%this treats genera within the same fam, from the same source as separate sources
  left_join(total_obs) %>%
  mutate(freq_squared = (n_family/n)^2) %>% 
  group_by(scientificName,n) #%>%
  summarize(degree_family = n_distinct(plant_family), 
            degree5_family = sum(n_family >= 5), weighted_degree_fam = sum(freq_squared),
            simpson_family_citations=sum(n_source_citations),simpson())

@@

globi_degree = degree_by_fam %>% left_join(degree_by_genus) %>%
  mutate(diet_breadth = ifelse(scientificName %in% specialists,'specialist','generalist'))
# write_csv(globi_degree,'modelign_data/globi_degree.csv')
